\documentclass[JohnsonMADraft2.tex]{subfiles}
\begin{document}
\section{Coding Notes}\label{CodingNotes}
\begin{itemize}
	\item Arkansas- In 1868, the Chief Justice was appointed by the Governor, with Senate Consent, with the Associate Justices elected by the people.  This is coded as partisan election since the majority of the justices are elected.
	
	\item Connecticut- from 1784 through 1818 Judges were appointed by the Legislature, but the tenure in office was dependent upon the action of the legislature.  These years are coded as missing and subsequently dropped from the dataset.
	
	\item Delaware- The original court of last resort was the Court of Appeals, and the data is coded as such.  The Supreme Court was created by constitutional amendment in 1951, at which point the data is coded to reflect the Supreme Court.
	
	\item Michigan- In 1939, a constitutional amendment passed calling for non-partisan elections for judges, except the Supreme Court, which would continue to be nominated at party conventions.  This is coded as a non-partisan following practice in this field.
	
	\item New Jersey- AJS Data does not reflect the Constitutional amendment in 1983, which extended subsequent terms to a term of good behavior.
	
	\item New Mexico- Judges are elected for the remainder of the unexpired term.  This could be up to 8 years, so it was coded as such.
	
	\item Oklahoma is examined individually, but both courts have mandatory jurisdiction.-2004
	
	\item Pennsylvania- From 1874 to 1968 Supreme Court justices were elected to twenty-one year terms and were not eligible for reelection.  This is coded as a 0 for the subsequent term.
	
	\item Tennessee- Constitution says qualified electors, however, by executive order the governor has created a nominating commission
	
	\item Texas is examined individually, SC has discretionary jurisdiction, but COA has mandatory in sentencing issues, so this is coded as mixed.- 2004
	
	\item Retention Elections- Gubernatorial reappointment, judicial commission reappointment, legislative reappointment are coded as No Retention Elections.
	
	\item Docket Control- This is established by looking at Criminal and Civil cases.  Administrative agency decisions are omitted, as well as death penalty cases.
	
	\item Docket Control- Observations between BJS reports are assumed to be static.
\end{itemize}	


\end{document}