\documentclass[JohnsonMADraft2]{subfiles}
\begin{document}
I have specified three individual types of models.  A primary specification with four indicators and 200 years of data, an alternative specification with five indicators and 42 years of data, and a regime change specification consisting of regime change observations over 200 years with four indicators.

Each model was run in R using Martyn \citeauthor{rjags}'s JAGS package and the CODA package for convergence diagnostics \cite{R,CODA}.\footnote{Other software packages used are listed in the Reference section without being cited in-text.}\nocite{R,CODA,R-Foreign,R2jags,ggplot2,dplyr,rjags}  Each model was run using 4 chains with 15,000 draws per chain.  The first 10,000 draws were thrown away as burn-in.

\subsection*{Primary Specification}
The primary model is composed of all state year observations between 1800 and 2012.  There are four indicators in this model: Appointment Method, Initial Term Length, Subsequent Term Length and Retention Method.  There are \todo{List number of Observations} observations.  

\subsection*{Alternative Specification}


\subsection*{Regime Change Specification}
The final model specification consists of the four original indicators, but is limited to only unique observations of each judicial institution rather than repeated observations over long periods of time.  This specification dramatically reduces the observations in the real data, but 

%\singlespacing
%\bibliographystyle{apsr}
%\bibliography{measurementbib}	
\end{document}
