\documentclass[12pt]{article}
\usepackage{setspace}
\usepackage{mathtools}
\usepackage{nicefrac}
\usepackage{fullpage}
\usepackage{times}
\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{listings}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{lscape}
\usepackage{hyperref}
\usepackage[affil-it]{authblk}
\usepackage{color}
%setup hyperlinks color
\definecolor{PSU}{RGB}{0,0,153}
\hypersetup{
	colorlinks=true,       % false: boxed links; true: colored links
	linkcolor=black,          % color of internal links (change box color with linkbordercolor)
	citecolor=PSU,        % color of links to bibliography
	filecolor=PSU,      % color of file links
	urlcolor=PSU           % color of external links
}

\title{\textit{Reassessing Random Assignment on the U.S. Courts of Appeals}\\ \vspace{5mm}Summer Research Proposal}

\author{\href{http://www.jeremyrjohnson.org/}{Jeremy R.\ Johnson}\\ \href{mailto:Jeremy.Johnson@psu.edu}{Jeremy.Johnson@psu.edu}}
\affil{Pennsylvania State University}
\date{\today}



\begin{document}
\maketitle
\thispagestyle{empty}
\pagebreak

\singlespacing
\setcounter{page}{1}
\pagenumbering{gobble}  %This turns off page numbers  

\subsection*{Introduction}
Due to an increasing interest in assessing causality, many recent articles have leveraged randomized panel and case assignment on the U.S. Court of Appeals to learn about judicial decision making and racial and gender characteristics of Court of Appeals judges \citep{Kastellec2010,Glynn2015,Farhang2014}.  The United States Courts of Appeals are broken into eleven geographic circuits plus the D.C. Circuit and the Federal Circuit.  These courts' primary responsibility is to hear appeals from the 94 U.S. District Courts.  Traditionally these courts hear cases in panels of three judges.  In normal practice, panels are randomly chosen then a series of cases are randomly assigned to these panels \citep{Hooper2011, Journalist2011,Chilton2014,Songer2007}.%\footnote{There are obvious exceptions to the random assignment such as judge availability, or conflicts of interest.}  
These assignments are made to all current circuit judges and available senior judges.  There is variation from circuit to circuit on panel length and the number of cases that each panel hears before judges are assigned to a different panel as well as the number of judges on the circuit; however, the randomness is nominally standard across all circuit courts.  

While this process is generally taken to be the normal procedure in the Federal Appellate courts, a substantial literature has challenged this as fact \citep{Atkins1974,Brown2000,Chilton2014}.  \citet{Chilton2014} performed Monte Carlo simulations and concluded that the observed partisan breakdown of panel assignments for a five year period could not have happened using random assignment.  \citet{Atkins1974} as well as \citet{Brown2000} examine ``panel packing'' during the 1960's.  In addition, \citet{Brown2000} examines modern practices qualitatively to determine if such practices could still go on today, determining that the courts have enough rule making discretion to do so.

\citet{Chilton2014} have shown that the panel assignment process is not entirely random.  If it is not entirely random, what then can help us predict how panels are assigned?  The most likely hypothesis for answering this question surrounds partisanship, both of the judges on the panel, as well as the chief judge, who exercises administrative control over the circuit.  As an extension of that administrative control, what role does the Circuit Executive and Circuit Clerk play in assignments?  By integrating existing data with a simulation-based approach to assess randomness, we are able to examine the influences of a host of judge- and court-level factors on the extent to which, relative to our expectation, judges are over- or under-represented on Court of Appeals panels. Moreover, we can do so while examining nearly a century's worth of assignments, allowing us to assess changes over time in patterns of nonrandom assignment.

\subsection*{Data}
The data for this project will come from the Songer Court of Appeals Database for the case data \citep{Songer2007}.  The data for the individual judge level variables will come from the Federal Judicial Center biographical database with supplemental information coming from the U.S. Court of Appeals Judge Attribute Database \citep{FJC,Gryski2008}.  Measures of case salience are already coded into the \citeauthor{Songer2007} Dataset which reduces any additional coding, using the method developed by \citet{Hettinger2003}.

\subsection*{Analysis}
The analysis of this data will proceed in two phases.  The first phase will be to create a simulated dataset of panel assignment probabilities.  This will consist of creating a set of available judges for each circuit/year.  This can be compared to the caseload statistics available to establish a baseline.  The \citeauthor{Songer2007} database is a stratified sample of all Court of Appeals cases from 1925-2002, which is approximately 20,000 case observations.  The simulation will establish a baseline of panel and case assignment probability which will show how under or over represented a judge is on the panels.  We will then apply a hierarchical generalized linear model with various judge level and circuit level effects.  

\subsection*{Tasks to Be Performed / Timeline}
The most important task will be to assemble the dataset and clean the data for analysis.  Both the \citeauthor*{Songer2007} and Federal Judiciary Biographical datasets are readily available, but will need to be cleaned and prepared.  The next task will be to match the judge and circuit attributes to the cases in the \citeauthor{Songer2007} dataset.  This will be completed by Jeremy over the course of June.  All cases will be coded for all case, judge, and circuit level variables, which should also be completed by the end of June.  Concurrently with gathering the data will be to build the simulated panel assignment dataset.  This will also be completed by the end of or mid-July.  The next step will be to build the model which will be done by Chris and Jeremy and should be completed by mid-July.  The final task should be the completion of the manuscript, which will be completed by Chris and Jeremy and should be done during the first week of August.  This article should be ready for submission by the beginning of August.

\singlespacing
\scriptsize
\bibliographystyle{apsr}
\bibliography{panelassbib}

\end{document}
